---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}Factor Investing and Asset Pricing Anomalies}
   \rfoot{\color{headergrey}Chapter 8}
   \lfoot{\color{headergrey}\thepage}
   \fancyfoot[C]{\rmfamily\color{headergrey}Machine Learning for Factor Investing}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

knitr::opts_chunk$set(
   echo = T, 
   eval = TRUE, 
   dev = 'png', 
   fig.width = 9, 
   fig.height = 3.5)

options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)

library(quantmod, quietly = TRUE, warn.conflicts = FALSE)
library(tidyquant, quietly = TRUE, warn.conflicts = FALSE)
library(PerformanceAnalytics, quietly = TRUE, warn.conflicts = FALSE)
library(xtable, quietly = TRUE, warn.conflicts = FALSE)

library(here, quietly = TRUE, warn.conflicts = FALSE)

library(tidyverse, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

data_dir <- here::here("Machine Learning for Factor Investing")
```

```{r pander_setup, include = FALSE}
knitr::opts_chunk$set(comment = NA)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
```

## Datasets

```{r}
load(file.path(data_dir, "data_ml.RData"))
```

### Example Factor, Size

```{r}

data_ml %>%
   group_by(date) %>% # group by date
   mutate(large = Mkt_Cap_12M_Usd > median(Mkt_Cap_12M_Usd)) %>% # Creates the cap sort
   ungroup() %>% # ungroup
   mutate(year = lubridate::year(date)) %>% # Creates a gear variable
   group_by(year, large) %>% # Analyze by year & cap
   summarize(avg_return = mean(R1M_Usd)) %>% # avg return by year & cap
   ggplot(aes(x = year, y = avg_return, fill = large)) + # plot!
   geom_col(position = "dodge") + # bars side-to-side
   theme(legend.position = c(0.8, 0.2)) + # legend location
   coord_fixed(124) + # x/y aspect ration
   theme(legend.title = element_blank()) +
   scale_fill_manual(values = c("#F87E1F", "#0570EA"), name = "", # colors
                     labels = c("Small", "Large")) +
   ylab("Average Returns") +
   theme(legend.text = element_text(size=9))


   
```

## Factors

### Size

__SMB__ = small firms minus large firms

### Value

__HM__ = high minus low: undervalued minus 'growth' firms

### Momentum

__WML__ winners minus losers

### Profitability

__RMW__ = robust minus weak profits

profitability is measured as (revenues - (cost and expenses)) / equity

### Investment

__CMA__ conservative minus aggressive

Investment is measured via the growth of total assets (divided by total assets).

### Low 'risk'

__BAB__ betting against beta

(simple vol, market beta, idiosyncratic vol, etc)

[Kenneth French Factor Library](https://mba.tuck.dartmouth.edu/pages/faculty/ken.french/data_library.html)

#### Example Factor Model

```{r}

min_date <- "1963-07-31"; max_date <- "2020-06-30"

temp <- tempfile()

KF_website <- "http://mba.tuck.dartmouth.edu/pages/faculty/ken.french/"
KF_file <- "ftp/F-F_Research_Data_5_Factors_2x3_CSV.zip"

link <- paste0(KF_website, KF_file)

download.file(link, temp, quiet = T)

FF_factors <- read_csv(unz(temp, "F-F_Research_Data_5_Factors_2x3.CSV"), 
                       skip = 3) %>%          # Check the number of lines to skip!
    rename(date = X1, MKT_RF = `Mkt-RF`) %>%  # Change the name of first columns
    mutate_at(vars(-date), as.numeric) %>%                 # Convert values to number
    mutate(date = ymd(parse_date_time(date, "%Y%m"))) %>%  # Date in right format
    mutate(date = rollback(date + months(1)))              # End of month date

FF_factors <- FF_factors %>% mutate(MKT_RF = MKT_RF / 100,
                                    SMB = SMB / 100,
                                    HML = HML / 100,
                                    RMW = RMW / 100,
                                    CMA = CMA / 100,
                                    RF = RF / 100) %>%
   filter(date >= min_date, date <= max_date)

knitr::kable(head(FF_factors), booktabs = T,
             caption = "Sample of Monthly Factor Returns.")

```

```{r}
FF_Avg_Returns <- FF_factors %>%
   mutate(date = year(date)) %>%
   gather(key = factor, value = value, - date) %>%
   group_by(date, factor) %>%
   summarise(value = mean(value))
```

```{r, fig.cap="Factor Returns"}

FF_Avg_Returns %>%
   ggplot(aes(x = date, y = value, color = factor)) +
   geom_line() + coord_fixed(500)

```

```{r, fig.cap="Return Densitites"}

FF_factors %>%
   gather(key = factor, value = return, - date) %>%
   filter(factor != 'RF') %>%
   ggplot(aes(return, group = factor)) +
      geom_density(aes(fill = factor, alpha = .25))

```


```{r, fig.cap="Yearly Avg Return Densities"}

FF_Avg_Returns %>%
   filter(factor != 'RF') %>%
   ggplot(aes(value, group = factor)) +
      geom_density(aes(fill = factor, alpha = .25))

```

```{r, fig.cap="Growth of $1 by factor"}

FF_Cum_Returns <- FF_factors %>%
   gather(key = factor, value = value, -date) %>%
   group_by(factor) %>%
   mutate(lag_ret = lag(value)) %>%
   mutate(return = cumprod(1 + ifelse(is.na(lag_ret), 0, lag_ret)))

FF_Cum_Returns %>%
   ggplot(aes(date, return, group = factor)) +
   
   geom_line(aes(col = factor))
```

## Fama-Macbeth Regressions

```{r}
separation_date <- as_date("2014-01-15")

traning_sample <- filter(data_ml, date < separation_date)
testing_sample <- filter(data_ml, date > separation_date)

stock_ids <- levels(as.factor(data_ml$stock_id)) # list of all stock ids

stock_days <- data_ml %>%
   group_by(stock_id) %>%
   summarise(nb = n())

stock_ids_short <- stock_ids[which(stock_days$nb == max(stock_days$nb))] # keep only stocks with full data
```

```{r}
nb_factors <- 5                                                     # Number of factors
data_FM <- left_join(data_ml %>%                                    # Join the 2 datasets
                         dplyr::select(date, stock_id, R1M_Usd) %>% # (with returns...
                         filter(stock_id %in% stock_ids_short),     # ... over some stocks)
                     FF_factors, 
                     by = "date") %>% 
    group_by(stock_id) %>%                                          # Grouping
    mutate(R1M_Usd = lag(R1M_Usd)) %>%                              # Lag returns
    ungroup() %>%
    na.omit() %>%                                                   # Remove missing points
    spread(key = stock_id, value = R1M_Usd)

models <- lapply(paste0("`", stock_ids_short, 
                        '` ~  MKT_RF + SMB + HML + RMW + CMA'),           # Model spec
                 function(f){ lm(as.formula(f), data = data_FM,           # Call lm(.)
                                 na.action="na.exclude") %>%       
                         summary() %>%                                    # Gather the output
                         "$"(coef) %>%                                    # Keep only coefs
                         data.frame() %>%                                 # Convert to dataframe
                         dplyr::select(Estimate)}                         # Keep the estimates
                 )
betas <- matrix(unlist(models), ncol = nb_factors + 1, byrow = T) %>%     # Extract the betas
    data.frame(row.names = stock_ids_short)                               # Format: row names
colnames(betas) <- c("Constant", "MKT_RF", "SMB", "HML", "RMW", "CMA")    # Format: col names

```

```{r}
knitr::kable(head(betas), caption = "Coefficents")
```

```{r}

loadings <- betas %>%
   dplyr::select(-Constant) %>%
   data.frame()

returns <- data_FM %>%
   dplyr::select(-MKT_RF, -SMB, -HML, -RMW, -CMA)

ret <- returns %>%
   dplyr::select(-date) %>%
   data.frame(row.names = returns$date) %>%
   t()

FM_data <- cbind(loadings, ret)
```

```{r}

models <- lapply(paste("`", 
                       returns$date, "`", 
                       ' ~  MKT_RF + SMB + HML + RMW + CMA', sep = ""),
                           function(f){ lm(as.formula(f), data = FM_data) %>%                        # Call lm(.)
                              summary() %>%                                    # Gather the output
                              "$"(coef) %>%                                    # Keep only the coefs
                              data.frame() %>%                                 # Convert to dataframe
                              dplyr::select(Estimate)}                         # Keep only estimates
                 )

gammas <- matrix(unlist(models), ncol = nb_factors + 1, byrow = T) %>%    # Switch to dataframe
    data.frame(row.names = returns$date)                                  # & set row names
colnames(gammas) <- c("Constant", "MKT_RF", "SMB", "HML", "RMW", "CMA")   # Set col names

```

```{r, fig.cap="Gammas"}

gammas %>%                                         # Take gammas:
    # The first row is omitted because the first row of returns is undefined
    dplyr::select(MKT_RF, SMB, HML) %>%                             # Select 3 factors
    bind_cols(date = data_FM$date) %>%                              # Add date
    gather(key = factor, value = gamma, -date) %>%                  # Put in tidy shape
    ggplot(aes(x = date, y = gamma, color = factor)) +              # Plot
    geom_line() + facet_grid( factor~. ) +                          # Lines & facets
    scale_color_manual(values=c("#F87E1F", "#0570EA", "#F81F40")) + # Colors
    coord_fixed(980)                                                # Fix x/y ratio
```

