---
title: ''
mainfont: Arial
fontsize: 12pt
documentclass: report
header-includes:
- \PassOptionsToPackage{table}{xcolor}
- \usepackage{caption}
- \usepackage{amssymb}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage[table]{xcolor}
- \usepackage{fancyhdr}
- \usepackage{boldline}
- \usepackage{tipa}
   \definecolor{headergrey}{HTML}{545454}
   \definecolor{msdblue}{HTML}{1C93D1}
   \pagestyle{fancy}
   \setlength\headheight{30pt}
   \rhead{\color{headergrey}\today}
   \fancyhead[L]{\color{headergrey}Moretz, Brandon}
   \fancyhead[C]{\Large\bfseries\color{headergrey}An Abridged Introduction to Finance}
   \rfoot{\color{headergrey}\thepage}
   \lfoot{\color{headergrey}Chapter 1}
   \fancyfoot[C]{\rmfamily\color{headergrey}Computational Finance}
geometry: left = 1cm, right = 1cm, top = 2cm, bottom = 3cm
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
editor_options: 
  chunk_output_type: console
---


```{r knitr_setup, include = FALSE}

# DO NOT ADD OR REVISE CODE HERE
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, dev = 'png')
options(knitr.table.format = "latex")

```

```{r report_setup, message = FALSE, warning = FALSE, include = FALSE}

library(data.table, quietly = TRUE, warn.conflicts = FALSE)

assignInNamespace("cedta.pkgEvalsUserCode", c(data.table:::cedta.pkgEvalsUserCode, "rtvs"), "data.table")

library(ggplot2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(ggthemes, quietly = TRUE, warn.conflicts = FALSE)
library(knitr, quietly = TRUE, warn.conflicts = FALSE)
library(kableExtra, quietly = TRUE, warn.conflicts = FALSE)
library(Rblpapi, quietly = TRUE, warn.conflicts = FALSE)
library(scales, quietly = TRUE, warn.conflicts = FALSE)
library(pander, quietly = TRUE, warn.conflicts = FALSE)
library(dplyr, quietly = TRUE, warn.conflicts = FALSE)
library(formattable, quietly = TRUE, warn.conflicts = FALSE)
library(grid, quietly = TRUE, warn.conflicts = FALSE)
library(gridExtra, quietly = TRUE, warn.conflicts = FALSE)
library(png, quietly = TRUE, warn.conflicts = FALSE)
library(extrafont, quietly = TRUE, warn.conflicts = FALSE)
library(tinytex, quietly = TRUE, warn.conflicts = FALSE)
library(stringr, quietly = TRUE, warn.conflicts = FALSE)
library(lubridate, quietly = TRUE, warn.conflicts = FALSE)
library(reshape2, quietly = TRUE, warn.conflicts = FALSE)
library(ggrepel, quietly = TRUE, warn.conflicts = FALSE)
library(mnormt, quietly = TRUE, warn.conflicts = FALSE)
library(Ecdat, quietly = TRUE, warn.conflicts = FALSE)
library(MASS, quietly = TRUE, warn.conflicts = FALSE)
library(copula, quietly = TRUE, warn.conflicts = FALSE)
library(fGarch, quietly = TRUE, warn.conflicts = FALSE)
library(forecast, quietly = TRUE, warn.conflicts = FALSE)
library(tseries, quietly = TRUE, warn.conflicts = FALSE)
library(gmodels, quietly = TRUE, warn.conflicts = FALSE)
library(rugarch, quietly = TRUE, warn.conflicts = FALSE)
library(quantmod, quietly = TRUE, warn.conflicts = FALSE)

options(tinytex.verbose = TRUE)
suppressMessages(library("tidyverse"))

pretty_kable <- function(data, title, dig = 2) {
  kable(data, caption = title, digits = dig) %>%
    kable_styling(bootstrap_options = c("striped", "hover")) %>%
      kableExtra::kable_styling(latex_options = "hold_position")
}

theme_set(theme_light())

# Theme Overrides
theme_update(axis.text.x = element_text(size = 10),
             axis.text.y = element_text(size = 10),
             plot.title = element_text(hjust = 0.5, size = 16, face = "bold", color = "darkgreen"),
             axis.title = element_text(face = "bold", size = 12, colour = "steelblue4"),
             plot.subtitle = element_text(face = "bold", size = 8, colour = "darkred"),
             legend.title = element_text(size = 12, color = "darkred", face = "bold"),
             legend.position = "right", legend.title.align=0.5,
             panel.border = element_rect(linetype = "solid", 
                                         colour = "lightgray"), 
             plot.margin = unit(c( 0.1, 0.1, 0.1, 0.1), "inches"))

data.dir <- "D:/Projects/Statistical-Computing/datasets/"

setwd("D:/Projects/Statistical-Computing/RDS")

```

```{r pander_setup, include = FALSE}

knitr::opts_chunk$set(comment = NA)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

```

#### Chapter 1

### Ex 1.1

We shall use the __getSymbols__ function in the quantmod package to retrieve financial data for General Electric (__GE__).

```{r echo = T}

getSymbols("GE", src = "yahoo", from = "2000-01-01", to = "2009-12-30")

names(GE)

GE["2000-01-01/2000-01-20"]

geAdj = GE$GE.Adjusted["2000-01-01-/2000-01-20"]; geAdj

max(geAdj); min(geAdj); mean(geAdj)

```

```{r echo = T, fig.height=3.5, fig.width=8}
chartSeries(GE)
```

```{r echo = T, fig.height=3.5, fig.width=8}
chartSeries(GE, TA=NULL, subset='2001-01::2001-02')
```

```{r echo=T}
saveRDS(GE, file = "GE.rds")
```

### 1.3.4

```{r, echo = T}

symbols <- c('^VLIC', 'GE', 'KO', 'AAPL', 'MCD')
getSymbols( symbols, src = "yahoo", from = "2012-02-01", to = "2013-02-01")

# obtain Adjusted Close

VLICad <- VLIC$VLIC.Adjusted; GEad <- GE$GE.Adjusted
KOad <- KO$KO.Adjusted; AAPLad <- AAPL$AAPL.Adjusted
MCDad <- MCD$MCD.Adjusted

# compute cumulative sum (cumsum) of daily returns (Delt)

vl <- cumsum( (Delt(VLICad) * 100)[-1, ])
ge <- cumsum( (Delt(GEad) * 100)[-1, ])
ko <- cumsum( (Delt(KOad) * 100)[-1, ])
ap <- cumsum( (Delt(AAPLad) * 100)[-1, ])
md <- cumsum( (Delt(MCDad) * 100)[-1, ])

### Range for the plot

lim <- c(min(vl, ge, ko, ap, md), max(vl, ge, ko, ap, md))

```

```{r, echo = T, fig.width=8, fig.height=4}
### plot

stocks <- data.table( Date = index(vl),
                      VLIC = as.numeric(vl), 
                      GE = as.numeric(ge), 
                      KO = as.numeric(ko), 
                      AAPL = as.numeric(ap), 
                      MCD = as.numeric(md))

stocks.long <- melt(stocks, 
     id.vars = c("Date"), 
     variable.name = c("Stock"), 
     value.name = "Return")

ggplot(stocks.long, aes(Date, Return, col = Stock)) +
   geom_line() +
   labs(title = "Stock Performance")

```

### 1.3.6

Using no arbitrage arguments show that for options on stocks:

i.) the stock's price is an upper bound for the price of a call;

ii.) the strike price is an upper bound for the price of a put.

### 1.3.7

The following is a list of well-known investment strategies obtained by different combinations of put and call options on the same underlying asset. For each one of these strategies compute the payoff function and draw the profit graph.

Additionally, argue about the situations where the strategy is profitable.

```{r, echo = T}

plot.payoff <- function(payoff, strike) {
   
   breakeven <- payoff[min(which(payoff$Payoff > 0))]$Strike
   ggplot(payoff) +
      geom_point(aes(Strike, Payoff), col = ifelse(payoff$Payoff < 0, "darkred", "darkgreen"), lwd = .8) +
      geom_hline(yintercept = 0, col = "cornflowerblue", lwd = 1, alpha = .45) +
      labs(title = paste0("Option Payoff: K=$", strike, ", Break-Even: $", breakeven), 
           x = "Strike", y = "Payoff") +
      scale_x_continuous(labels = dollar) +
      scale_y_continuous(labels = dollar)
}

```

#### Vanilla Call Payoff

```{r, echo = T, fig.height=3.5, fig.width=8}

call.payoff <- function( strike, initial.price, contracts = 1, contract.size = 100,
                         lower = -.15, upper = .15) {

   cost <- -(initial.price * contract.size) * contracts
   
   x <- strike + seq(strike*lower, strike*upper, .01)
   y <- ( ( x - strike - initial.price ) * contract.size ) * contracts

   payoff <- ifelse(y < cost, cost, y)
   
   data.table(Strike = x, Payoff = payoff)
}

K <- 45
plot.payoff(call.payoff(K, 2.88), K)

```

#### Vanilla Put Payoff

```{r, echo = T, fig.height=3.5, fig.width=8}

put.payoff <- function( strike, initial.price, contracts = 1, contract.size = 100,
                         lower = -.15, upper = .15) {

   cost <- -(initial.price * contract.size) * contracts
   
   x <- strike + seq(strike*lower, strike*upper, .01)
   y <- ( ( strike - x - initial.price ) * contract.size ) * contracts

   payoff <- ifelse(y < cost, cost, y)
   
   data.table(Strike = x, Payoff = payoff)
}

K <- 45
plot.payoff(put.payoff(K, 2.45), K)

```

#### The Straddle:

The straddle strategy consists of two legs, one long and one short, both with the same strike and expiration date.

The main idea behind the straddle is that it is a non-directional long volatility strategy. It is generally suitable when you expect the underlying security to be very volatile and move a lot, but you are not sure whether the price move will be up or down. The position makes a profit when your expectation is correct and the underlying does make a big move to one side or the other. If the underlying price stays more or less the same, the trade makes a loss.

Straddle Payoff:

```{r, echo = T, fig.height=3.5, fig.width=8}

K <- 45
plot.payoff(call.payoff(K, 2.85) + put.payoff(K, 2.88), K)

```

#### The strangle:

Similar to the straddle. It consists on buying a put and a call with the same expiration date, but different strike prices. If $K_c$ is the strike price for the call, and $K_p$ is the strike price for the put, then the strategy requires $K_c > K_p$.

```{r, echo = T}

```

#### The strip:

This strategy consists of long positions in one call and two puts, all with the same strike price and expiration date.

```{r, echo = T}

```

#### The strap:

This one consists of long positions in two calls and one put, all with the same strike price and expiration date.

```{r, echo = T}

```

#### The butterfly spread:

This is made with options of the same type. Suppose we use calls and athe underlying assset is a stock. Then a butterfly spread of calls consists on short selling two calls waith strike price $K_0$ close to the current stock price, and buying two calls, one with strike price $K_0-c$ and the other with strike price $K_0 + c$, where 2c > 0 is the length of the spread chosen by the investor.

```{r, echo = T}

```